\documentclass[12pt]{article}

\usepackage[cm]{fullpage}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{fancyhdr}
\usepackage{titling}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{enumitem}


\begin{document}
\thispagestyle{empty}


\section{Exposé}

\subsection{Title}
TODO

\subsection{Problem}

The configuration management library Elektra supports the mounting of different file types, like ini, json or xml, in order to handle configuration. With Elektra, it's possible to read/write these configuration files within one interface, making developers or administrators more independent of the used filetypes. However, it isn't possible to mount TOML configuration files in Elektra yet.

\subsection{Goal/Question}
The goal of this thesis is to implement a storage plugin to enable the usage of TOML files with Elektra.
To support as many platforms as possible, the plugin will be written purely in C. Parsing will be done with flex and bison.

Afterwards, a case study will be conducted on the OSS project lcdproc, which is currently undergoing electrification.
In the case study, the newly implemented TOML plugin will be used in lcdproc for handling of configuration.
It will be checked, if the plugin integrates well within the application and can load the needed configuration information.
Additionally, it will be measured if the plugin imposes any unreasonable overhead on loading times.
For this, it will be compared against other existing storage plugins.

\subsection{Method}

To answer the research question, benchmarks will be used.

\subsection{Timeplan/Milestones}
\begin{enumerate}
\item 12/10/2019: Storage Plugin finished, Thesis Introduction, Related Work
\item 22/10/2019: Benchmarking finished, Thesis Methodology, Part of Evaluation
\item 31/10/2019: Thesis Evaluation, Conclusion 
\end{enumerate}

\subsection{Related Work/References}
\begin{itemize}
\item lcdproc: http://lcdproc.omnipotent.net/
\item TOML: https://github.com/toml-lang/toml
\item TOML ABNF: https://github.com/toml-lang/toml/blob/master/toml.abnf 
\end{itemize}

\end{document}

